---
title: "practical_exercise_3, Methods 3, 2021, autumn semester"
author: 'Otto Sejrskild Santesson'
date: "5 October 2021"
output: html_document
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(utils,tidyverse, patchwork, ggplot2, lme4, stats, grid, ggpubr, ggrepel, graphics,effects,VCA, vroom, readbulk,stringi, gridExtra, MuMIn,dfoptim)
```

# Exercises and objectives
The objectives of the exercises of this assignment are:  
1) Download and organise the data and model and plot staircase responses based on fits of logistic functions  
2) Fit multilevel models for response times  
3) Fit multilevel models for count data  

REMEMBER: In your report, make sure to include code that can reproduce the answers requested in the exercises below (__MAKE A KNITTED VERSION__)  
REMEMBER: This assignment will be part of your final portfolio

## Exercise 1

Go to https://osf.io/ecxsj/files/ and download the files associated with Experiment 2 (there should be 29).  
The data is associated with Experiment 2 of the article at the following DOI https://doi.org/10.1016/j.concog.2019.03.007  


1) Put the data from all subjects into a single data frame  
2) Describe the data and construct extra variables from the existing variables  
    i. add a variable to the data frame and call it _correct_ (have it be a _logical_ variable). Assign a 1 to each row where the subject indicated the correct answer and a 0 to each row where the subject indicated the incorrect answer (__Hint:__ the variable _obj.resp_ indicates whether the subject answered "even", _e_ or "odd", _o_, and the variable _target_type_ indicates what was actually presented.
```{r}
df <- read_bulk(
  directory= "experiment2/",# "Audition/" is the foldername where data is stored
  fun = read_csv
)

df = df %>% 
  mutate(correct = ifelse(target.type == "odd" & obj.resp == "o",1,
                          ifelse(target.type == "even" & obj.resp == "e",1,0)))
```

ii. describe what the following variables in the data frame contain, _trial.type_, _pas_, _trial_, _target.contrast_, _cue_, _task_, _target_type_, _rt.subj_, _rt.obj_, _obj.resp_, _subject_ and _correct_. (That means you can ignore the rest of the variables in your description). For each of them, indicate and argue for what `class` they should be classified into, e.g. _factor_, _numeric_ etc.  

*_trial.type_*: a character variable encoded as either "staircase" (i.e. data was collected as a part of the adaptive staircase procedure which was performed at the beginning of the study) or "experiment" (i.e. data was collected as a part of the actual experiment trials). It should be re-encoded a factor in order for _trial.type_ to be modelled later in the analysis, if relevant. 
```{r}
df$trial.type <- as.factor(df$trial.type)
```

*_pas_*: indicates one of the 4 categorically different ratings of the Perceptual Awareness Scale (PAS: Ramsøy & Overgaard, 2004), rated by the participant. It indicates the participant's perceptual awareness of the stimulus where the numbers represent the following:* 

*(1) No experience* (i.e. "no impression of the stimulus. All answers are seen as mere guesses)*

*(2) Weak Glimpse*  (i.e. "A feeling that something has been shown. Not characterized by any content, and this cannot be specified any further)*

*(3) Almost Clear Experience* (i.e. "Ambiguous experience of the stimulus. Some stimulus aspects are experienced more vividly than others. A feeling of almost being certain about one’s answer")*

*4) Clear experience* (i.e. "Non-ambiguous experience of the stimulus. No doubt in one’s answer")*

It should be encoded as a factor as it is a case of ordinal data.
```{r}
df$pas <- as.factor(df$pas)
```

*_trial_*: refers to the trial number. It is numeric but should be factor
```{r}
df$trial <- as.factor(df$trial)
```

*_target.contrast_*: indicates the contrast of the target stimulus, relative to the background, adjusted to match the threshold of each individual by using the QUEST-algorithm. It is numeric and should stay numeric as we are dealing with a continuous variable with values falling within the range of 0-1*

*_cue_*: cue shown to participants to indicate the set of possible digits that might appear on each trail. There were 36 different combination of cues (0-35). The task-column specifies the number of potential targets cued. As the _cue_ variable contains nominal data, it should be reencoded into a factor.

```{r}
df$cue <- as.factor(df$cue)
```


*_task_*: indicates the number of potential targets that were cued. Consisted of either 2,4, or 8 digits, indicated by the character variable *_task_* taking one of the 3 values:

  *singles* (2 possible targets) e.g. 2:9
  *pairs* (4 possible targets) e.g. 24:57
  *quadruplet* (8 possible targets) e.g. 2468:3579

This variable should also be re-encoded as a factor as we are, again, dealing with nominal data.
```{r}

df$task <- as.factor(df$task)

```

*target_type*: indicates the parity of the target stimulus (i.e. whether the digit was even or odd). It is encoded as a character but should be re-encoded as a factor. 
```{r}
df$target.type <- as.factor(df$target.type)
```

*rt.subj*: response time of answering Perceptual Awareness Scale
Encoded as numeric as it is continuous data.

*rt.obj*: response time (time taken by participant to indicate the parity of the target stimulus). 
Encoded as numeric as it is continuous data.

*obj.resp*: the participant's answer when asked to indicate the parity of the target stimulus (e = even, o = odd). Encoded as numeric but should be re-encoded as a factor.
```{r}
df$obj.resp <- as.factor(df$obj.resp)
```

*subject*: participant ID, enabling us to distinguish between data collected from different participants. Is a character variable but should be re-encoded as a factor as we are dealing with a repeated measures design and therefore need to be able to model subject as random intercepts in the analysis phase.

```{r}
df$subject <- as.factor(df$subject)
```

*correct*: indicates accuracy of the respondent's answer of the target stimulus' parity. It is a binary variable where 0 indicates an incorrect response and 1 indicates a correct one, and should therefore be re-encoded as a factor. 

```{r}
df$correct <- as.factor(df$correct)
```



1) Put the data from all subjects into a single data frame  
2) Describe the data and construct extra variables from the existing variables  
    i. add a variable to the data frame and call it _correct_ (have it be a _logical_ variable). Assign a 1 to each row where the subject indicated the correct answer and a 0 to each row where the subject indicated the incorrect answer (__Hint:__ the variable _obj.resp_ indicates whether the subject answered "even", _e_ or "odd", _o_, and the variable _target_type_ indicates what was actually presented.
    ii. describe what the following variables in the data frame contain, _trial.type_, _pas_, _trial_, _target.contrast_, _cue_, _task_, _target_type_, _rt.subj_, _rt.obj_, _obj.resp_, _subject_ and _correct_. (That means you can ignore the rest of the variables in your description). For each of them, indicate and argue for what `class` they should be classified into, e.g. _factor_, _numeric_ etc.  
    
    
    
  iii. for the staircasing part __only__, create a plot for each subject where you plot the estimated function (on the _target.contrast_ range from 0-1) based on the fitted values of a model (use `glm`) that models _correct_ as dependent on _target.contrast_. These plots will be our _no-pooling_ model. Comment on the fits - do we have enough data to plot the logistic functions?  
  
```{r}
dfStair = df %>% 
  subset(trial.type == "staircase")

dfStair$subject = gsub("(?<![0-9])0+", "", dfStair$subject, perl = TRUE)
dfStair$subject = as.integer(dfStair$subject)

nopoolfun <- function(i){
  dat <- dfStair[which(dfStair$subject == i),]
  model <- glm(correct~target.contrast, family = 'binomial', data=dat)
  fitted <- model$fitted.values
  plot_dat <- data.frame(cbind(fitted,'target.contrast'=dat$target.contrast))
plot <- ggplot(plot_dat, aes(x = target.contrast, y = fitted))+
  geom_point()+
  geom_line(aes(x = target.contrast, y = fitted))+
  xlab('Target Contrast')+
  ylim(c(0,1))+
  theme_minimal()
return(plot)
}

subjects <- c(1:16)
plots <- lapply(subjects, FUN=nopoolfun)
do.call(grid.arrange,  plots)
subjects <- c(17:29)
plots <- lapply(subjects, FUN=nopoolfun)
do.call(grid.arrange,  plots)
```



```{r}
#  By looking at the different plots, the fitted values doesn't seem to follow the traditional sigmoid function of logistic regression - the s-formed curve, that indicates the probability of 1, given the predictor variable(s). We might therefore need more data 
Data enoug for it to seem like there is a 



# There seem to be "enough" data, since I can stil plo the logistic function (even though the typical sigmmoid function isn't ) #
```
  
iv. on top of those plots, add the estimated functions (on the _target.contrast_ range from 0-1) for each subject based on partial pooling model (use `glmer` from the package `lme4`) where unique intercepts and slopes for _target.contrast_ are modelled for each _subject_  
    v. in your own words, describe how the partial pooling model allows for a better fit for each subject  

```{r}
partialPoolModel = glmer(correct~target.contrast + (target.contrast|subject), family = 'binomial', dfStair)
dfStair$yEstPartialPooling = fitted(partialPoolModel)

poolNoPoolFun <- function(i){
  dat <- dfStair[which(dfStair$subject == i),]
  model <- glm(correct~target.contrast, family = 'binomial', data=dat)
  fitted <- model$fitted.values
  plot_dat <- data.frame(cbind(fitted,'target.contrast'=dat$target.contrast,'yEstPartialPooling' = dat$yEstPartialPooling))
plot <- ggplot(plot_dat)+
  geom_point(aes(x = target.contrast, y = fitted), color = "#3f00ff")+
  geom_line(aes(x = target.contrast, y = fitted), color = "#3f00ff")+
  geom_point(aes(x = target.contrast, y = yEstPartialPooling), color = "#d82e21")+
  geom_line(aes(x = target.contrast, y = yEstPartialPooling), color = "#d82e21")+
  xlab('Target Contrast')+
  ylim(c(0,1))+
  theme_minimal()
return(plot)
}


subjects <- c(1:16)
plots <- lapply(subjects, FUN=poolNoPoolFun)
do.call(grid.arrange,  plots)
subjects <- c(17:29)
plots <- lapply(subjects, FUN=poolNoPoolFun)
do.call(grid.arrange,  plots)
```
Beep boop - blue is the no pooling model, and red is the partial pooling model.
It is better at generalizing

## Exercise 2

Now we __only__ look at the _experiment_ trials (_trial.type_)  

1) Pick four subjects and plot their Quantile-Quantile (Q-Q) plots for the residuals of their objective response times (_rt.obj_) based on a model where only intercept is modelled  
    i. comment on these    
    ii. does a log-transformation of the response time data improve the Q-Q-plots?  
```{r}
dfExp = df %>% 
  subset(trial.type == "experiment")

dfExp4 = dfExp %>% 
  filter(subject == "004")

dfExp10 = dfExp %>% 
  filter(subject == "010")

dfExp17 = dfExp %>% 
  filter(subject == "017")

dfExp20 = dfExp %>% 
  filter(subject == "020")


dfExp4Model = lm(rt.obj ~ 1, data = dfExp4)
dfExp10Model = lm(rt.obj ~ 1, data = dfExp10)
dfExp17Model = lm(rt.obj ~ 1, data = dfExp17)
dfExp20Model = lm(rt.obj ~ 1, data = dfExp20)

qqnorm(resid(dfExp4Model));qqline(resid(dfExp4Model), col = "blue")
qqnorm(resid(dfExp10Model));qqline(resid(dfExp10Model), col = "green")
qqnorm(resid(dfExp17Model));qqline(resid(dfExp17Model), col = "red")
qqnorm(resid(dfExp20Model));qqline(resid(dfExp20Model), col = "purple")
```
Doing an visual inspection of the QQ-plots, the residuals don't seem to be normally distributed at all (and therefore breaking one of the assumption of linear regression) - the values from the higher quantiles of the residuals diverges relatively much from the straight line, that they're supposed to follow, if they were normally distributed.

We will therefore try to do a log-transformation of the response time data, to see if this will improve the QQplots:
```{r}
dfExp4ModelLog = lm(log(rt.obj) ~ 1, data = dfExp4)
dfExp10ModelLog = lm(log(rt.obj) ~ 1, data = dfExp10)
dfExp17ModelLog = lm(log(rt.obj) ~ 1, data = dfExp17)
dfExp20ModelLog = lm(log(rt.obj) ~ 1, data = dfExp20)

qqnorm(resid(dfExp4ModelLog));qqline(resid(dfExp4ModelLog), col = "blue")
qqnorm(resid(dfExp10ModelLog));qqline(resid(dfExp10ModelLog), col = "green")
qqnorm(resid(dfExp17ModelLog));qqline(resid(dfExp17ModelLog), col = "red")
qqnorm(resid(dfExp20ModelLog));qqline(resid(dfExp20ModelLog), col = "purple")
```
Looking at the QQPlots of the log-transformed data, they look way better now - all QQplots, except the one of the log transformed data of subject 10, more or less fits the line of the QQplot, which signifies that the residuals of the log-transformed data is more normally distributed compared to the non-log-transformed data


2) Now do a partial pooling model modelling objective response times as dependent on _task_? (set `REML=FALSE` in your `lmer`-specification)  
    i. which would you include among your random effects and why? (support your choices with relevant measures, taking into account variance explained and number of parameters going into the modelling)  
    ii. explain in your own words what your chosen models says about response times between the different tasks  
    
```{r}
parPoolRtTaskModel1 = lmer(rt.obj ~ task + (1|subject), dfExp, REML = FALSE)

parPoolRtTaskModel2 = lmer(rt.obj ~ task +(1|trial), dfExp, REML = FALSE)
#Convergence error - signifies an overfit and therefore the model is hard tot generalize to new data

parPoolRtTaskModel3 = lmer(rt.obj ~ task + (1|subject) + (1|trial), dfExp, REML = FALSE)

parPoolRtTaskModel4 = lmer(rt.obj ~ task + (task|subject) + (1|trial), dfExp, REML = FALSE)
# Convergence error - signifies an overfit and therefore the model is hard tot generalize to new data

AIC(parPoolRtTaskModel1,parPoolRtTaskModel2,parPoolRtTaskModel3,parPoolRtTaskModel4)

sigma(parPoolRtTaskModel1) # 2.858998
sigma(parPoolRtTaskModel2) # 2.878982
sigma(parPoolRtTaskModel3) # 2.858263
sigma(parPoolRtTaskModel4) # 2.857295

r.squaredGLMM(parPoolRtTaskModel1) #R2c = 0.014649
r.squaredGLMM(parPoolRtTaskModel2) #R2c = 0.00082565
r.squaredGLMM(parPoolRtTaskModel3) #R2c = 0.015158
r.squaredGLMM(parPoolRtTaskModel4) #R2c = 0.015826
```

    
3) Now add _pas_ and its interaction with _task_ to the fixed effects  
    i. how many types of group intercepts (random effects) can you add without ending up with convergence issues or singular fits?  
    ii. create a model by adding random intercepts (without modelling slopes) that results in a singular fit - then use `print(VarCorr(<your.model>), comp='Variance')` to inspect the variance vector - explain why the fit is singular (Hint: read the first paragraph under details in the help for `isSingular`)
    iii. in your own words - how could you explain why your model would result in a singular fit?  
```{r}
convergTestModel1 = lmer(rt.obj ~ task + pas:task + (1|subject), dfExp, REML = FALSE)
convergTestModel2 = lmer(rt.obj ~ task + pas:task + (1|subject) + (1|trial), dfExp, REML = FALSE)
convergTestModel3 = lmer(rt.obj ~ task + pas:task + (1|subject) + (1|trial) + (1|cue), dfExp, REML = FALSE)

convergTestModel4 = lmer(rt.obj ~ task + pas:task + (1|subject) + (1|trial) + (1|cue) + (1|target.type), dfExp, REML = FALSE)

convergTestModel5 = lmer(rt.obj ~ task + pas:task + (1|subject) + (1|trial) + (1|cue) + (1|target.type) + (1|odd.digit), dfExp, REML = FALSE)

convergTestModel6 = lmer(rt.obj ~ task + pas:task + (1|subject) + (1|trial) + (1|cue) + (1|target.type) + (1|odd.digit) + (1|even.digit), dfExp, REML = FALSE)

print(VarCorr(convergTestModel6), comp='Variance')
```
The amount of group intercepts than can be added does not only depend on the amount, but also on the chosen group intercepts - but in our case, with the group intercepts chosen, it took 6 group intercepts before we ended up with singular fits

"ii.By adding (1|subject)+(1|trial)+(1|task) to the existing model with "pas" and "task" as fixed effects interaction, the model met singular fits issues. When inspecting the variance vector of the model with singular fits (np_model_4), it shows that the reason for the singular fit, is that the model contains a random-effect variance estimate of zero, this random-effect is (1|task)."

"iii. I suppose the singular fit is a result of overfitting. The model and its random effects are to complex to be supported by the data. To overcome this, we could remove some random effects one by one."


## Exercise 3

1) Initialize a new data frame, `data.count`. _count_ should indicate the number of times they categorized their experience as _pas_ 1-4 for each _task_. I.e. the data frame would have for subject 1: for task:singles, pas1 was used #times, pas2 was used # times, pas3 was used # times and pas4 was used # times. You would then do the same for task:pairs and task:quadruplet  

```{r}
data_count <- df %>% 
  group_by(subject, task, pas) %>% 
  dplyr::summarize("count" = n())
view(data_count)

#EXTRA STUFF

# Double-checking that there are no observations that fits with the following specifications, since it is not in the data frame of 'data_count':
df %>% filter(subject == "002",task == "pairs",pas == 4)

# We see, that there are no observations that meet the above specifications (the output of the above line gives a datafram with 0 rows), and therefore the row specifying this has been left out. If we want the rows that includes the different combinations of subject, pas and task, that has 0 as the amount of observations, we need to construct the data frame differently or mutate the data frame by inserting the rows (since I haven't found an easier way to do it and for the fun of it).
# Since the data frame is going to be used for fitting regression lines, which makes it important to include all groups, even though where count = 0, we'll try constructing the data frame differently:

baseDf = data.frame(pas = rep(1:4,(3*29)),
                    task = rep(c(rep("singles",4),rep("pairs",4),rep("quadruplet",4)),29),
                    subject = rep(1:29,each = (3*4)))

df_copy = df
df_copy$subject = gsub("(?<![0-9])0+", "", df_copy$subject, perl = TRUE)
df_copy$subject = as.integer(df_copy$subject)

count = c()

for (i in 1:nrow(baseDf)){
countVector = df_copy %>%  
  filter(subject == baseDf[i,"subject"],task == baseDf[i,"task"],pas == baseDf[i,"pas"]) %>% 
  nrow()
count = append(count,countVector)
}

dataCount = cbind(count,baseDf)
dataCount$pas = as.factor(dataCount$pas)
dataCount$task = as.factor(dataCount$task)
dataCount$subject = as.factor(dataCount$subject)
```


```{r}

# To check which family we should use, we make a histogram of the count variable to visually inspect the distribution of the data. 
hist(dataCount$count)
# We see, that the curve follows the one of a poisson distribution, which makes sense, since we're dealing with data that specifies counts - the data points therefore always have a positive x-value. Furthermore, the x-values are all integers.

# We now make the model, specifying the family to be poisson, based on the nature of the data's distribution (no intercept for subject is being modeled - only the slope for the variable 'pas' for each subject):
count_model = glmer(count ~ pas*task + (pas+0|subject), data = dataCount, family = poisson, control = glmerControl(optimizer = "bobyqa")) 
summary(count_model)

count_model2 = glmer(count ~ pas*task + (pas+0|subject), data = dataCount, family = poisson, control = glmerControl(optimizer = "bobyqa")) 
summary(count_model);summary(count_model2)

```

2) Now fit a multilevel model that models a unique "slope" for _pas_ for each _subject_ with the interaction between _pas_ and _task_ and their main effects being modelled  
    i. which family should be used?  
    ii. why is a slope for _pas_ not really being modelled?  
    
  Question and Lau's answer:
  
  Question: Question 3.2.ii.: I don't know what you mean with "why is a slope for _pas_ not really being modelled?"

    - Lau: try to have a look at the output of the random effects (ranef) or the design matrix Zt using getME  
    
  iii. if you get a convergence error, try another algorithm (the default is the _Nelder_Mead_) - try (_bobyqa_) for which the `dfoptim` package is needed. In `glmer`, you can add the following for the `control` argument: `glmerControl(optimizer="bobyqa")` (if you are interested, also have a look at the function `allFit`)
    iv. when you have a converging fit - fit a model with only the main effects of _pas_ and _task_. Compare this with the model that also includes the interaction  
    v. indicate which of the two models, you would choose and why  
    vi. based on your chosen model - write a short report on what this says about the distribution of ratings as dependent on _pas_ and _task_  
    vii. include a plot that shows the estimated amount of ratings for four subjects of your choosing 
3) Finally, fit a multilevel model that models _correct_ as dependent on _task_ with a unique intercept for each _subject_  
    i. does _task_ explain performance?  
    ii. add _pas_ as a main effect on top of _task_ - what are the consequences of that?  
    iii. now fit a multilevel model that models _correct_ as dependent on _pas_ with a unique intercept for each _subject_
    iv. finally, fit a model that models the interaction between _task_ and _pas_  and their main effects  
    v. describe in your words which model is the best in explaining the variance in accuracy  

